### 2015-05-11 intro.tex
update in intro

Data races lead to nondeterministically occurring bugs that can be hard to reproduce, isolate and debug, especially in the context of a complex operating system. According to the C language standard, data races also lead to undefined behavior, which can be the source of various compilation and optimization issues \PDComment{which is the best cite here?}. Linux kernel analyzers, such as sparse~\cite{corbet2004sparse}, coccinelle~\cite{padioleau2008doc} and lockdep~\cite{corbet2006lock}, have successfully found deadlocks in kernel code, but are typically unable to detect data races. Techniques such as~\cite{ball2006thorough, clarke2004predicate, qadeer2004kiss, engler2000checking, henzinger2002temporal, cook2006termination, pratikakis2006locksmith, kuznetsov2010testing, renzelmann2012symdrive, lal2012corral} have been used to analyze the program behavior of Linux and Windows device drivers, but primarily focus on reasoning about generic sequential program properties and finding protocol bugs. The few of these techniques that attempt to deal with data races are usually either \emph{unsound} (i.e. can miss real bugs) or \emph{imprecise} (i.e. can report false bugs). Furthermore, most previous techniques that attempt to reason about thread interleavings face significant scalability issues, because of the exponentially large state-space of realistic concurrent programs~\cite{musuvathi2008finding}.

@@

We then show that by exploiting the race-freedom guarantees that our symbolic analysis provides, we can achieve a sound partial-order reduction that significantly accelerates \corral~\cite{lal2012solver}, a state-of-the-art bug-finder that is used by Microsoft to analyze device drivers, as well as other concurrent programs. Using \whoop and \corral we analyzed \sizeOfBenchmarks drivers available in the 4.0 Linux kernel distribution.

>>>>

Data races lead to nondeterministically occurring bugs that can be hard to reproduce, isolate and debug, especially in the context of complex operating systems. Further, according to the C language standard, races result in undefined behavior, which can be the source of tricky compilation and optimization issues (as compilers typically assume race-freedom). Realistic concurrent programs also have an exponentially large state-space, which presents a serious challenge for implementing efficient race-checking techniques~\cite{musuvathi2008finding}.

Techniques such as~\cite{ball2006thorough, clarke2004predicate, engler2000checking, henzinger2002temporal, cook2006termination, kuznetsov2010testing, renzelmann2012symdrive, lal2012corral} have been successfully used to analyze device drivers, but mostly focus on generic sequential program properties and finding protocol bugs. Well-regarded Linux kernel analyzers, such as sparse~\cite{corbet2004sparse}, coccinelle~\cite{padioleau2008doc} and lockdep~\cite{corbet2006lock}, can find deadlocks in kernel source code, but are unable to detect races. Techniques that can detect races in drivers~\cite{qadeer2004kiss, pratikakis2006locksmith, voung2007relay, lal2012corral} are usually either \emph{unsound} (i.e. can miss real bugs) or \emph{imprecise} (i.e. can report false bugs), and typically sacrifice precision for scalability. Thus, there is a clear need for new tools that are able to detect races efficiently and precisely.

@@

We then show that by exploiting the race-freedom guarantees that our symbolic analysis provides, we can achieve a sound partial-order reduction that significantly accelerates \corral~\cite{lal2012corral}, a state-of-the-art bug-finder that is used by Microsoft to analyze device drivers, as well as other concurrent programs. Using \whoop and \corral we analyzed \sizeOfBenchmarks drivers available in the 4.0 Linux kernel distribution.

### 2015-05-11 related.tex
some fixes
update in intro

Using static analysis to detect data races is a promising alternative to dynamic concurrency bug finding, which is inherently limited by scalability and execution path coverage issues. Notable previous work includes the static analyzers Warlock~\cite{sterling1993warlock}, ESC~\cite{detlefs1998extended} and LockLint~\cite{oracle2010locklint}. However, all these three tools heavily rely on user annotations, and thus require a lot of manual effort that limits their applicability. \textsc{Whoop}, on the other hand, does not require any source code modifications, and thus can be applied with zero effort.

Most related to our work are the static lock set analyzers RELAY~\cite{voung2007relay} and Locksmith~\cite{pratikakis2006locksmith}. Both tools, though, have significant limitations. The authors of RELAY found 5022 warnings when analyzing the Linux kernel, with only 25 of them being true data races. To limit these false positives, RELAY employs post-analysis unsound filters, but these can also filter out true races. Although Locksmith successfully detected data races in several small Pthreads applications and 7 medium-sized Linux device drivers, it also reported a significant number of false positives. The authors also reported that Locksmith was unable to run on several large programs, showcasing its limited scalability.

>>>>

Statically analyzing concurrent programs to detect data races is a promising alternative to dynamic techniques, which commonly face scalability and code coverage issues as they typically rely on a (controlled) scheduler to explore execution paths. Notable previous work includes the static analyzers Warlock~\cite{sterling1993warlock}, ESC~\cite{detlefs1998extended} and LockLint~\cite{oracle2010locklint}. However, all these three tools heavily rely on user annotations, and thus require a lot of manual effort that limits their applicability. \textsc{Whoop}, on the other hand, does not require any source code modifications, and thus can be applied with zero effort.

Most related to our work are the static lockset analyzers RELAY~\cite{voung2007relay} and Locksmith~\cite{pratikakis2006locksmith}. Both tools, though, have significant limitations. The authors of RELAY found 5022 warnings when analyzing the Linux kernel, with only 25 of them being true data races. To limit these false positives, RELAY employs post-analysis unsound filters, but these can also filter out true races. Although Locksmith successfully detected data races in several small Pthreads applications and 7 medium-sized Linux device drivers, it also reported a significant number of false positives. The authors also reported that Locksmith was unable to run on several large programs, showcasing its limited scalability.
